---
title: "r511 Readme"
author: "Tom Buckley"
date: "7/9/2018"
output: 
  github_document:
    toc: true
---

[![Travis-CI Build Status](https://travis-ci.com/r-gtfs/r511.svg?branch=master)](https://travis-ci.com/r-gtfs/r511)
[![cran version](https://www.r-pkg.org/badges/version/r511)](https://cran.r-project.org/package=r511)

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = FALSE,
  comment = "#>",
  message = FALSE,
  error = FALSE,
  warning = FALSE,
  fig.path = "README/README-",
  fig.width=7.3,
  fig.height=5,
  out.width = '100%'
)

knitr::opts_chunk$set(echo = TRUE)
library(r511)
```

## Set your api key as an environmental variable

If you don't have a key, you can get one here:  
https://511.org/developers/list/tokens/create

```{r}
#Sys.setenv(APIKEY511="yourkeyhere")
```

## Get MTC 511 Operator List

This function pulls a list of operator names, modes, and private codes. The latter we use to make requests for GTFS data. 

```{r}
operator_df <- get_511_metadata()
head(operator_df)
```

## Get URL for GTFS Data

You can use the get_511_url() function to build a URL from which you can directly download GTFS data for an operator. 

```{r}
bart_code <- operator_df[operator_df$name=='Bay Area Rapid Transit',]$privatecode
bart_gtfs_url <- get_511_url(bart_code)
```

## Import Data

Using [trread](https://github.com/r-gtfs/trread), load BART data into R as a list of dataframes. 

```{r}
library(trread)
bart_gtfs_data <- import_gtfs(bart_gtfs_url)
```

## Example: Summarise Stops Per Route

Summarise the number of stops per route on BART. 

```{r, message=FALSE, warning=FALSE}
library(dplyr)
attach(bart_gtfs_data)

routes_df %>% inner_join(trips_df, by="route_id") %>%
  inner_join(stop_times_df, by="trip_id") %>% 
    inner_join(stops_df, by="stop_id") %>% 
      group_by(route_long_name) %>%
        summarise(stop_count=n_distinct(stop_id)) %>%
  arrange(desc(stop_count))
```


